---
title: CV
name: Genevieve
surname: Chiong
position: " "
address: Daly City, CA
phone: +1 (415) 215-4622
www: mariecurie.com
email: "gchiong@ucsb.edu"
github: genchiong
linkedin: genevievechiong
date: "`r format(Sys.time(), '%B %Y')`"
aboutme: "hi! "
output: 
  vitae::markdowncv:
    theme: kjhealy
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

## Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Master of Environmental Science and Management", "Expected 2022", "Bren School of Environmental Science & Management - University of California, Santa Barbara", "Santa Barbara",
  "Bachelor of Science in Ecology, Behavior, and Evolution", "2014", "University of California, Los Angeles", "Los Angeles",
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

## Project Management and Community Engagement Experience

```{r}
tribble(
  ~Year, ~Type, ~Desc,
  1903, "Physics", "Awarded for her work on radioactivity with Pierre Curie and Henri Becquerel",
  1911, "Chemistry", "Awarded for the discovery of radium and polonium"
) %>% 
  brief_entries(
    glue::glue("Nobel Prize in {Type}"),
    Year, 
    Desc
  )
```

## Communication and Public Speaking Experience

```{r consulting}
tribble(
  ~ role, ~ company, ~ loc, ~ dates, ~ details,
  "Data scientist", "Nectric", "Clayton, Australia", "Jan. 2019 - Present", "A variety of consulting projects with substantial spatio-temporal modelling tasks.",
  "Research assistant", "Monash University", "Clayton, Australia", "Jan. 2016 - Present", "Involved in many internal and external data projects of varying size and complexity.",
  "Research assistant", "Monash University", "Clayton, Australia", "Jan. 2016 - Present", "Projects typically feature large scale time series modelling or analysis, cross-sectional modelling and creating tools for improving workflows.",
  "Research assistant", "Monash University", "Clayton, Australia", "Jan. 2016 - Present", "Consulting project clients include: Huawei, Monash University, R Consortium, NSW Chief Scientist office, DiabetesLab and Tennis Australia.",
  "Data mining intern", "iSelect", "Cheltenham, Australia", "Feb. 2015 - Mar. 2015", "Improved business reporting with interactive visualisations, and model-based anomaly detection.",
) %>% 
  detailed_entries(role, dates, company, loc, details)
```

```{r}
tribble(
  ~Year, ~Type, ~Desc,
  1903, "Physics", "Awarded for her work on radioactivity with Pierre Curie and Henri Becquerel",
  1911, "Chemistry", "Awarded for the discovery of radium and polonium"
) %>% 
  brief_entries(
    glue::glue("Nobel Prize in {Type}"),
    Year, 
    Desc
  )
```

## Additional Field and Lab Experience

```{r}
tribble(
  ~Year, ~Type, ~Desc,
  1903, "Physics", "Awarded for her work on radioactivity with Pierre Curie and Henri Becquerel",
  1911, "Chemistry", "Awarded for the discovery of radium and polonium"
) %>% 
  brief_entries(
    glue::glue("Nobel Prize in {Type}"),
    Year, 
    Desc
  )
```

## Skills and Certifications

**Computer**: R, QGIS, Microsoft Office Suite, Google Suite, Salesforce, Zoom, GoToWebinar, SurveyMonkey, Mailchimp, Canva, Photoshop
**Languages**: English (Fluent), Mandarin (Conversational)
Certifications: Accounting Fundamentals (City College of San Francisco)
**Awards**: University of San Francisco Whitehead Research Fellowship ($10,000)
**Presentations**: Presented acoel Convolutriloba longifissura project to 100+ people at the University of San Francisco Creative Active and Research Day 2017

