---
title: "Text Analysis: The Red Headed League by Sir Arthur Conan Doyle"
description: |
  A short description of the post.
author:
  - name: Genevieve
    url: {}
date: 02-24-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(textdata)
library(ggwordcloud)
library(readtext)
library(pdftools)
library(tidytext)
```

```{r}
# Read in the data
red_headed_league_text <- pdf_text("red_headed_league.pdf")
```

```{r}
# Wrangle data to get tokens into tidy format, removing stop words
red_headed_league_tidy <- data.frame(red_headed_league_text) %>% 
  mutate(text_full = str_split(red_headed_league_text, pattern = "\\n")) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full)) %>% 
  unnest_tokens(word, text_full) %>% 
  dplyr::select(-red_headed_league_text) %>% 
  anti_join(stop_words)
```

```{r}
# Create a subset of the top 150 words in the whole text 
top_150 <- red_headed_league_tidy %>% 
  count(word) %>% 
  arrange(-n) %>% 
  slice(1:150)

# Word cloud of counts for the most frequently used words in the entire text
top_150_cloud <- ggplot(data = top_150,
                        aes(label = word)) +
  geom_text_wordcloud(aes(color = n, 
                          size = n), 
                      shape = "star") +
  scale_size_area(max_size = 8) +
    scale_color_gradientn(colors = c("goldenrod3", "violetred", "steelblue4")) +
  theme_minimal()

top_150_cloud
```

**Figure 1.** The 150 most frequently used words in Sir Author Conan Doyle's "The Red Headed League". Data: Doyle, A.C. 1892.

```{r}
# Perform sentiment analysis using the "nrc" lexicon on the entire text
red_headed_league_nrc <- red_headed_league_tidy %>% 
  inner_join(get_sentiments("nrc"))

red_headed_league_nrc_counts <- red_headed_league_nrc %>% 
  count(sentiment)

# Create a final visualization
red_headed_league_nrc_counts %>% 
  mutate(sentiment = fct_reorder(sentiment, n)) %>% 
  ggplot(aes(x = sentiment,
             y = n,
             fill = sentiment)) +
  geom_col() +
  coord_flip() +
  theme_minimal() +
  labs(x = "Sentiment\n",
       y = "\nWord Count") +
  theme(legend.position = "none")

```

**Figure 2.** Sentiment analysis of Sir Arthur Conan Doyle's "The Red Headed League" using the NRC Emotion Lexicon. Data: Doyle, A.C. 1892.; Mohammed, S. and Turney, P. 2013.)


**Citations**

Doyle, A.C. 1892. The Adventures of Sherlock Holmes London, England: George Newnes Ltd. https://etc.usf.edu/lit2go/32/the-adventures-of-sherlock-holmes/346/adventure-2-the-red-headed-league/

Crowdsourcing a Word-Emotion Association Lexicon, Saif Mohammad and Peter Turney, Computational Intelligence, 29 (3), 436-465, 2013.

